Visualization and analysis of probability distributions of large temporal data

### 2022-02-24 report/response.Rmd

I thank my examiners for their thorough and constructive comments. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

# Professor Catherine Hurley

## Chapter 1

> This would be a good place to detail the co-authorship arrangements for the succeeding chapters, and for the student’s contribution to be stated clearly.

@@

> Q10. In table 3.1 $N_c$ is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
3.3.3) it is m.

Done.

Add $m$ to table 3.1
$m$ = number of cyclic granularities to display together

$N_c$ refers to the total number of contextual cyclic granularities. $m$ refers to the number of cyclic granularities we are considering together in the display. For example, contextual cyclic granularities could be $hour_{day}$, $day_{wee}k$ and $month_{year}$ and we want to visualize any one granularity at a time. So $N_c = 3$ and $m=1$.

@@

> Q17. In the 3.3.1 algorithm, there is m, and then M.

Changed.
1.b) for $i \in {1, 2, \dots, M}$.

> Q18. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.

$m=1$ removed from the main chapter and included in supplements.

@@

> Q21. In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
should refer to the rug. Maybe show fewer panels?

Done.

> Q22. In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend. Maybe use a different colour in the heatmap for the significant comparisons. State the threshold for significance in the caption. In Table 3.3 the caption should explain the threshold. (Maybe use colour instead of stars to indicate significance?)

CHANGED.
- Different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border 
- threshold explained in caption in Table 3.3 
- grey color explained in the text and caption

CLARIFICATION 1: Using color to convey significance instead of stars (in the Fig 3.3a) is a not a great idea because the varied shades of the tiles represent different values of the $wpd$. For example, both tiles may be significant, and their colours may differ to represent distinct $wpd$ values. To be consistent, Table 3.3 also has stars to represent significance.

@@

What is the question? I have mentioned that D stands for Jensen-Shannon distances.

> Q8. In Table 4.1, the R column should have 250, not 20.

Done.

> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the distribution of each $v_t$.

Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables.

@@

Done.

> Q17. Again, the computational burden could do with more Discussion. At present, if I were to use this method on my data, what kind of sizes are realistic to work with?

To do.(Hard)

Suppose there are $n$ observations. Our current methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs. The time taken to compute distances between a pair of observations based on $p$ cyclic granularity is

@@

# Prof Juergen Symanzik

## Chapter 1

> Q1. While the author hints at other sources of similar data, I think it would be worthwhile to give the reader some further specific ideas here in the Introduction where similar complex data may occur, e.g., data from traffic sensors, movement data (turnstiles in metro stations or public buildings), and even more traditional measurements such as temperatures or precipitation over time that could benefit from the methods and software described in this thesis.

Done. Added in the Introduction.

> Q2. It may also be helpful to state what is not be covered in this thesis: The spatial component of such data where an additional component could be latitude and longitude or some areal code, such as the post code or some administrative units. Extending the work from this thesis with spatial proximity methods for clustering (https://doi.org/10.1016/j.cageo.2011.12.017) [assuming that households in certain residential neighborhoods may have similar energy usage patterns] and trying spatial visualization methods (if possible) such as glyph maps (https:
//onlinelibrary.wiley.com/doi/abs/10.1002/env.2152) should be mentioned somewhere as a consideration for future work but that may become a future new thesis by itself.

The name of the thesis implies that only analyzing temporal data is covered. Extending it to spatial is too broad a scope. The limitations of the study cater to aspects that could be addressed with some more time. This is not the case for extending the spatial component, which, as accurately pointed out, would require time comparable to another thesis.

@@

> Q16. Match the text on p. 50 and introduce the abbreviations from Figure 3.7 and Table 3.3, e.g., "hod" likely is "hour day" and so on. Introducing these abbreviations on p. 53 comes
too late.

Done.
Added the abbreviations on Page 50 in "Choosing cyclic granularities of interest and removing clashes" section.

> Q17. In Figure 3.7a, it is not clear what the eight heatmaps represent. Do they belong to id 1, ..., id 8 ? Some explanation in the text or listing the id information on the left would be helpful. Similarly, on p. 53, the text states: id 7 and 8 have the same significant harmonies. I do not know where to look at the heatmaps in Figure 3.7a to find this information (and what is mentioned under items 2. and 3. in the text). Some explanation and labeling is needed here.

Caption changed to illustrate the plot better. Added box borders in (a) such that (a) and (b) display the household ids in the same order as indicated by the same color of the line plot in (b) and box border in (a).

### 2022-02-24 Rmd/03-hakear.Rmd
Q17 (Juergen) and Q22(Caatherine) Ch3

(ref:dotplot-8) An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways. Panel (b) shows the raw demand series for September to highlight the repetitive patterns of energy demand. Panel (a) shows $\wpd$ values across harmonies where a darker color indicates a higher ranking harmony. A significant harmony is shown with an asterisk. For example, ids 7 and 8 have significant patterns across (*hod*, *dow*) and (*dow*, *hod*). Panel (c) is useful for comparing households across harmonies. For example, for the harmony (*dow*-*hod*), ids 1 and 7 have the least and highest $\wpd$ respectively.

>>>>

(ref:dotplot-8) An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways. (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border in (a). (b) shows the raw demand series for September to highlight the repetitive patterns of energy demand. (a) shows $\wpd$ values across harmonies where a darker shade of red indicates a higher ranking harmony and gray represents a clash. A significant harmony is shown with an asterisk. For example, ids 7 and 8 have significant patterns across the pair (*hod*, *dow*) since the tile with *hod* as facet and *dow* as x variable has an asterisk. (c) is useful for comparing households across harmonies. For example, for the harmony (*dow*-*hod*), ids 1 and 7 have the least and highest $\wpd$ respectively.
